---
title: "project_187"
author: "Huaye Yang"
date: "4/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.width=8, 
                      fig.height=5) 
# Load the packages
pacman::p_load(gapminder, dplyr,tidyverse, skimr, socviz, grid, stringr,
               usmap, maps, statebins, viridis, leaflet)

```
### How were measurements taken, or questions asked? Do you suspect 
### any bias in the questions or measurements?

It looks like the data was gathered through survey because all the variables 
are categorical. We assume it was taken online (there are StartDate and EndDate 
that were recorded in minutes), and participants only need to check boxes that 
they think best describes their situation.

Bias can easily occur in this survey because:
1. It may be hard for participants to find a certain categorical box within 
each question to best describe their situation. In this circumstances, samples
that are considered as outlier will fall into one of the categorical values. 
Also, we are only able to do a rough analysis based on designed categorical 
values. It can't provide detailed information or quantitative analysis for 
accurate numerical values. 

2. Participants may overly/underly describe their situation due to the form of 
survey. What we have in dataset is subjective, not objective.

3. We are not sure whether this survey is voluntary. If this is forced or hard
for participants to refuse, the data can be messed because participants may 
randomly select the boxes instead of actually looking into the questions and 
participate. Also, we can't guarantee how reliable the data is. From the 
recorded time (EndDate - StartDate),participants only spent around 3 minutes
for answering 29 questions (Some of them only use less than 1 minute). There 
might be part of the data that are not meaningful.

If this survey is voluntary (or maybe taken at a counselling institution,etc),
participants who are willing to answer these questions may already have issues
or at least went through the situation that were asked in the survey. This 
means we are not getting randomly distributed samples.


```{r clean_data}

# Do data clean-up here!
# Read in data
sleeping_alone_data1 <- read.csv("C:/Users/huaye/Desktop/R_sleeping_alone.csv", stringsAsFactors = T)
# Take a look at the relationship status column, 7 categories
summary(sleeping_alone_data1)
# Rename columns to simpler names based on survey questions
sleeping_alone_data <- sleeping_alone_data1 %>%
  
  rename( Relationship_Status = Which.of.the.following.best.describes.your.current.relationship.status., 
         Relationship_Length = How.long.have.you.been.in.your.current.relationship..If.you.are.not.currently.in.a.relationship..please.answer.according.to.your.last.relationship.,
         How_Often_Sep_Beds = When.both.you.and.your.partner.are.at.home..how.often.do.you.sleep.in.separate.beds., 
         Diff_Bed_Where_You_Sleep = When.you.re.not.sleeping.in.the.same.bed.as.your.partner..where.do.you.typically.sleep., 
         Diff_Bed_Where_Partner_Sleep = When.you.re.not.sleeping.in.the.same.bed..where.does.your.partner.typically.sleep., 
         Sep_Beds_Reasons = What.are.the.reasons.that.you.sleep.in.separate.beds..Please.select.all.that.apply., 
         When_First_Time_Sep_Beds = When.was.the.first.time.you.slept.in.separate.beds., 
         Statement_Help_Stay_Together =  To.what.extent.do.you.agree.with.the.following.statement...sleeping.in.separate.beds.helps.us.to.stay.together.., 
         Statement_Better_Sleep =  To.what.extent.do.you.agree.with.the.following.statement...we.sleep.better.when.we.sleep.in.separate.beds..,
         Statement_Improved_Sex_Life = To.what.extent.do.you.agree.with.the.following.statement.éš..our.sex.life.has.improved.as.a.result.of.sleeping.in.separate.beds..éš., 
         Current_Occupation =  Which.of.the.following.best.describes.your.current.occupation.,
         Household_Income = Household.Income,  
         Location_Census_Region = Location..Census.Region.)
  
sleeping_alone_data <- sleeping_alone_data %>% slice(-c(1))
head(sleeping_alone_data)
```

### 2.2 Occupation and Income (Graph 2 and Graph 3)

# Relate occupation,gender and bed status with how often they sleep separately.

### 2.2.1 Graph 2
```{r 2b_graph2}
# change the default theme below:
theme_set(theme_bw())


# set levels in order to give a better visisualization in color
summary(sleeping_alone_data$How_Often_Sep_Beds)

rank<-as.factor(sleeping_alone_data$How_Often_Sep_Beds)
sleeping_alone_data <- sleeping_alone_data %>%
  mutate(How_Often_Sep_Beds = factor(rank, ordered = TRUE, 
                                     evels = c("Never",
                                               "Once a year or less",
                                               "Once a month or less",
                                               "A few times per month",
                                               "A few times per week",
                                               "Every night")))



# Relate occupation to bed status (and how often they sleep)
# Overall plot
sleeping_alone_data %>%
  group_by(Current_Occupation)%>%
  ggplot(data = sleeping_alone_data,
       mapping = aes(x = Current_Occupation,
                     fill = How_Often_Sep_Beds)) +

  geom_bar(position="dodge")+
  
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.25,
                                   size = 7),
        legend.position = "top",
        legend.text = element_text(size = 7),
        legend.key.size = unit(0.25,'cm'),
        legend.title = element_text(size= 8),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())


# summary() to see each category name and ready to draw plots for each of 
# the categories. facet_wrap or facet_grid provide too little information
summary(sleeping_alone_data$How_Often_Sep_Beds)

#### for each bed status below:

# Never
sleeping_alone_data %>%
  select(Current_Occupation,How_Often_Sep_Beds) %>%
  filter(How_Often_Sep_Beds == "Never") %>%
  ggplot(aes(y = Current_Occupation,
             color = How_Often_Sep_Beds))+
  geom_bar(fill = "purple4")+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())
  

# Once a year or less 
sleeping_alone_data %>%
  select(Current_Occupation,How_Often_Sep_Beds) %>%
  filter(How_Often_Sep_Beds == "Once a year or less") %>%
  ggplot(aes(y = Current_Occupation,
             color = How_Often_Sep_Beds))+
  geom_bar(fill = "darkslateblue")+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())



# Once a month or less
sleeping_alone_data %>%
  select(Current_Occupation,How_Often_Sep_Beds) %>%
  filter(How_Often_Sep_Beds == "Once a month or less") %>%
  ggplot(aes(y = Current_Occupation,
             color = How_Often_Sep_Beds))+
  geom_bar(fill = "cadetblue4")+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())



# A few times per month 
sleeping_alone_data %>%
  select(Current_Occupation,How_Often_Sep_Beds) %>%
  filter(How_Often_Sep_Beds == "A few times per month") %>%
  ggplot(aes(y = Current_Occupation,
             color = How_Often_Sep_Beds))+
  geom_bar(fill = "lightseagreen")+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())


# A few times per week
sleeping_alone_data %>%
  select(Current_Occupation,How_Often_Sep_Beds) %>%
  filter(How_Often_Sep_Beds == "A few times per week") %>%
  ggplot(aes(y = Current_Occupation,
             color = How_Often_Sep_Beds))+
  geom_bar(fill = "palegreen3")+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())



# Every night
sleeping_alone_data %>%
  select(Current_Occupation,How_Often_Sep_Beds) %>%
  filter(How_Often_Sep_Beds == "Every night") %>%
  ggplot(aes(y = Current_Occupation,
              color= How_Often_Sep_Beds))+
  geom_bar(fill = "khaki1")+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())


## for people who have further response instead of checking boxes
## (choose to plot those who answered "retired" or "Retired)

sleeping_alone_data %>%
  filter(X.12 == "retired" | X.12 == "Retired")%>%
  mutate(X.12 ="retired") %>%
  
  ggplot(aes(x = X.12, fill= How_Often_Sep_Beds)) +
  
  geom_bar(color="white",position = "dodge")+
  
  theme(axis.text.x = element_text(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) 


```
### 2.2.2 Summarizing Paragraph
People who are in Education, Training, and Library Occupations seems to have 
more response of "Never" sleeping separately. Regardless of the occupation,
The number of people who sleep separately is bigger than other frequencies.

For people who checked the box of "Other(please specify)", we select retired 
samples to draw a bar graph. Distribution of people who retired have heavy tails
on both side(Never & Every night), but the number of people who never sleeping 
separately is almost twice as large as those who sleeping separately every night.


### 2.2.3 Graph 3
```{r 2b_graph3}
# GRAPH 3
# Second graph could build on first (add income as a factor)
ggplot(data = sleeping_alone_data,
       mapping = aes(x = Current_Occupation,
                     fill = How_Often_Sep_Beds)) +
  geom_bar()+
  scale_fill_viridis_d(option = "cividis")+
  
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.25,
                                   size = 7),
        legend.position = "top",
        legend.text = element_text(size = 7),
        legend.key.size = unit(0.25,'cm'),
        legend.title = element_text(size= 8),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())+
  facet_grid(~Gender)

## for each bed status below:
# Never
sleeping_alone_data %>%
  select(Current_Occupation,How_Often_Sep_Beds,Gender) %>%
  filter(How_Often_Sep_Beds == "Never") %>%
  ggplot(aes(x= Current_Occupation,
             color = How_Often_Sep_Beds,
             fill = Gender))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.25,
                                   size = 7),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())+
  facet_grid(~Gender)
  

# Once a year or less 
sleeping_alone_data %>%
  select(Current_Occupation,How_Often_Sep_Beds,Gender) %>%
  filter(How_Often_Sep_Beds == "Once a year or less") %>%
  ggplot(aes(x= Current_Occupation,
             color = How_Often_Sep_Beds,
             fill = Gender))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.25,
                                   size = 7),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())+
  facet_grid(~Gender)

# Once a month or less
sleeping_alone_data %>%
  select(Current_Occupation,How_Often_Sep_Beds,Gender) %>%
  filter(How_Often_Sep_Beds == "Once a month or less") %>%
  ggplot(aes(x= Current_Occupation,
             color = How_Often_Sep_Beds,
             fill = Gender))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.25,
                                   size = 7),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())+
  facet_grid(~Gender)

# A few times per month 
sleeping_alone_data %>%
  select(Current_Occupation,How_Often_Sep_Beds,Gender) %>%
  filter(How_Often_Sep_Beds == "A few times per month") %>%
  ggplot(aes(x= Current_Occupation,
             color = How_Often_Sep_Beds,
             fill = Gender))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.25,
                                   size = 7),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())+
  facet_grid(~Gender)



# A few times per week
sleeping_alone_data %>%
  select(Current_Occupation,How_Often_Sep_Beds,Gender) %>%
  filter(How_Often_Sep_Beds == "A few times per week") %>%
  ggplot(aes(x= Current_Occupation,
             color = How_Often_Sep_Beds,
             fill = Gender))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.25,
                                   size = 7),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())+
  facet_grid(~Gender)



# Every night
sleeping_alone_data %>%
  select(Current_Occupation,How_Often_Sep_Beds,Gender) %>%
  filter(How_Often_Sep_Beds == "Every night") %>%
  ggplot(aes(x= Current_Occupation,
             color = How_Often_Sep_Beds,
             fill = Gender))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.25,
                                   size = 7),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())+
  facet_grid(~Gender)


```
### 2.2.4 Summarizing Paragraph:
When we add gender as another consideration of our plots, it seems that there 
are more male participants for this study. For each frequencies of how often 
they sleep separately, there is a significant difference in number for the 
response of Never & Every night(males more than females). For other response 
of this question, the number in difference is not significant. People who
does not specify their gender has a large proportion of responding to "Never".

