---
title: 'Stat/CS 187: Final Project'
author: "Huaye Yang, Halina Vercessi, and Matthew Mccarthy"
date: "Spring 2022"
output:
  word_document:
    toc: yes
  html_document:
    toc: yes
    df_print: paged
---
  

```{r setup, message = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.width=12, 
                      fig.height=10)

# Load the packages
pacman::p_load(gapminder, tidyverse, skimr, socviz, grid, stringr,
               usmap, maps, statebins, viridis, leaflet, cowplot, gridExtra)



```  

***

# 1. Introduction

Before this survey, the most detailed data and reports on this subject were 
gathered and made in 2004 by the National Sleep Foundation. The data from the 
National Sleep Foundation provided the numbers on how many couples slept separately 
but did not gather data that could offer insight as to the reasons couples sleep apart 
and why couples sleep in separate beds. This data is interesting because it is the 
kind of data that could be useful to psychologists and social workers. This data 
can be used to examine stereotypes we have about why couples sleep separately 
and which age ranges of couples sleep separately along with being useful for 
examining why couples sleep separately.The data comes from a survey created in 
2014 by Mona Chalabi as research for an article for Five Thirty Eight (from ABC News). 
The population from which this sample was drawn includes 1,057 American adults 
who were married, in a domestic partnership, in civil union or cohabiting with a 
significant other. The sample was selected with the help of SurveyMonkey, which 
gathered all 1,057 responses.  

The participants only need to check boxes that they think best describes their 
situation. Bias can easily occur in this survey because:

1. It may be hard for participants to find a certain categorical box within 
each question to best describe their situation. In this circumstances, samples
that are considered as outlier will fall into one of the categorical values. 
Also, we are only able to do a rough analysis based on designed categorical 
values. It can't provide detailed information or quantitative analysis for 
accurate numerical values. 

2. Participants may overly/underly describe their situation due to the form of 
survey. What we have in dataset is subjective, not objective.

3. We are not sure whether this survey is voluntary. If this is forced or hard
for participants to refuse, the data can be messed because participants may 
randomly select the boxes instead of actually looking into the questions and 
participate. Also, we can't guarantee how reliable the data is. From the 
recorded time (EndDate - StartDate),participants only spent around 3 minutes
for answering 29 questions (Some of them only use less than 1 minute). There 
might be part of the data that are not meaningful. If this survey is voluntary 
(or maybe taken at a counselling institution,etc), participants who are willing 
to answer these questions may already have issues or at least went through the 
situation that were asked in the survey. This means we are not getting randomly 
distributed samples.

The cleaning that was done on the data was to rename the columns because the 
original column names were built from the survey questions (separated by periods) 
and they were very long and tedious to work with. Additionally, during the building 
of the following graphs, responses that were left blank were omitted to focus on 
relevant responses. We also removed the first row as it served as almost a second title 
row but was being counted in the rows of data. We considered labeling the blank / 
no answer spaces as N/A but N/A means something in this data as it is a valid 
response to specific check boxes. 





```{r clean_data}

# Read in data
sleeping_alone_data1 <- read.csv("R_sleeping_alone.csv", stringsAsFactors = T)

# Take a look at the relationship status column, 7 categories
#summary(sleeping_alone_data1)

# Rename columns to simpler names based on survey questions
sleeping_alone_data <- sleeping_alone_data1 %>%
  
  rename( Relationship_Status = Which.of.the.following.best.describes.your.current.relationship.status., 
         Relationship_Length = How.long.have.you.been.in.your.current.relationship..If.you.are.not.currently.in.a.relationship..please.answer.according.to.your.last.relationship.,
         How_Often_Sep_Beds = When.both.you.and.your.partner.are.at.home..how.often.do.you.sleep.in.separate.beds., 
         Diff_Bed_Where_You_Sleep = When.you.re.not.sleeping.in.the.same.bed.as.your.partner..where.do.you.typically.sleep., 
         Diff_Bed_Where_Partner_Sleep = When.you.re.not.sleeping.in.the.same.bed..where.does.your.partner.typically.sleep., 
         Sep_Beds_Reasons = What.are.the.reasons.that.you.sleep.in.separate.beds..Please.select.all.that.apply., 
         When_First_Time_Sep_Beds = When.was.the.first.time.you.slept.in.separate.beds., 
         Statement_Help_Stay_Together =  To.what.extent.do.you.agree.with.the.following.statement...sleeping.in.separate.beds.helps.us.to.stay.together.., 
         Statement_Better_Sleep =  To.what.extent.do.you.agree.with.the.following.statement...we.sleep.better.when.we.sleep.in.separate.beds..,
         Statement_Improved_Sex_Life = To.what.extent.do.you.agree.with.the.following.statement.ë_.our.sex.life.has.improved.as.a.result.of.sleeping.in.separate.beds..ë_, 
         Current_Occupation =  Which.of.the.following.best.describes.your.current.occupation.,
         Household_Income = Household.Income,  
         Location_Census_Region = Location..Census.Region.)
  

sleeping_alone_data <- sleeping_alone_data %>% slice(-c(1))

head(sleeping_alone_data)

```

***

# 2. Data Visualizations


## 2.1 Length of Relationship and Age (Graph 1):  


**The following figures aim to elate the length of relationships, bed status (separate beds vs same bed), and age.**
**Additionally, the start of sleeping separately in looked at.**
**Is it more likely that couples who have been together longer will sleep in separate beds more often?**


### 2.1.1 Graph 1 --> Halina 
```{r 2a_graph1}

sleeping_alone_data2 <- sleeping_alone_data %>%
  
  filter(Relationship_Length != "", How_Often_Sep_Beds != "") %>%
  
  group_by(Relationship_Length, How_Often_Sep_Beds) %>%
  
  summarise(count_freq = n()) %>%
  
  mutate(Proportion = round(count_freq/sum(count_freq), digits = 5)) %>%
  
  select(Relationship_Length, How_Often_Sep_Beds, Proportion) %>% 
  
  slice(-c(1)) %>%
  
  ungroup()


rel_length_bar <- ggplot(data = sleeping_alone_data2,
                            mapping = aes(x = Relationship_Length,
                                          y = Proportion, 
                                          fill = How_Often_Sep_Beds)) +
  
                     geom_col(position = "dodge") + 
  
                     labs(x = "Relationship Length",
                          fill = "How Often Couples Sleep Separately", 
                          title = "Frequency Sleeping Separately and Relationship Length (proportions)")+
  
                    scale_y_continuous(labels = scales::percent) +
                    scale_x_discrete(name = "Relationship Length (years)", labels = c("1-5", "11-15", "16-20", "6-10","< 1", "> 20" )) +
                    facet_wrap(~How_Often_Sep_Beds) +
                    theme(legend.position = "NA", 
                          title = element_text(size = 9)) +
                    coord_flip()


```

```{r 2a_graph1.1}


sleeping_alone_data3 <- sleeping_alone_data %>%
  
  filter(Age != "", How_Often_Sep_Beds != "") %>%
  
  group_by(Age, How_Often_Sep_Beds) %>%
  
  summarise(count_freq = n()) %>%
  
  mutate(Proportion = round(count_freq/sum(count_freq), digits = 5)) %>%
  
  select(Age, How_Often_Sep_Beds, Proportion) %>% 
  
  slice(-c(1)) %>%
  
  ungroup()


prop_age_bar <- ggplot(data = sleeping_alone_data3,
                            mapping = aes(x = Age,
                                          y = Proportion, 
                                          fill = How_Often_Sep_Beds)) +
  
                     geom_col(position = "dodge") + 
  
                     labs(x = "Age",
                          fill = "How Often Couples Sleep Separately", 
                          title = "Frequency Sleeping Separately and Age (proportions)")+
  
                    scale_y_continuous(labels = scales::percent)+ 
                    scale_x_discrete(name = "Age Range", labels = c("60-Up", "18-29", "30-44", "45-60")) +
                    facet_wrap(~How_Often_Sep_Beds) +
                    theme(legend.position = "NA", 
                          title = element_text(size = 9)) +
                    coord_flip()


plot_grid(rel_length_bar, prop_age_bar, nrows = 2, ncols = 1) 

```



```{r 2a_graph1.2}

# Thinking about omitting this graph but we think it would be interesting to still examine the starting 
# point of sleeping separately within the relationship span. 

length_and_start_graph <- sleeping_alone_data %>%
  filter(Age != "", How_Often_Sep_Beds != "", When_First_Time_Sep_Beds != "") %>%
  ggplot(mapping = aes(x = Relationship_Length,
                      y = How_Often_Sep_Beds)) +

  geom_jitter(mapping = aes(color = When_First_Time_Sep_Beds)) +

  facet_wrap(facets = ~When_First_Time_Sep_Beds) +

  scale_x_discrete(name = "Relationship Length (years)", labels = c("1-5", "11-15", "16-20", "6-10","< 1", "> 20" )) +

  scale_y_discrete(name = "How Often Couple Sleeps in Separate Beds", labels = c("Few Times Per Month", "Few Times Per Week", "Every Night", "Never", "Once a Month or Less","Once a Year or Less")) +

  theme(legend.position = "NA") +

  labs(title = "Relationship Length, Frequency Sleeping in Separate Beds, and When it Started")

length_and_start_graph

```


### 2.1.2 Summarizing Paragraph:

From the first two graphs, it is clear that the majority of couples who took the 
survey do not sleep separately. From the first graph, around 18% of couples who
have been in a relationship for 16-20 years and more than 20 years do sleep in 
separate beds. These percentages are on the higher end for proportions of couples 
that have slept or do sleep separately. This aspect of the data fits our society's 
stereotyping that couples who have been together longer are more likely to sleep 
separately. Another notable percentage is that almost 20% of couples who have been 
together for less than a year reported sleeping separately once a month or less. 

From the last graph, which aims to show patterns in when sleeping separately starts, 
it appears that the majority of couples who do sleep in separate beds every night 
started sleeping separately earlier on in the relationship (within 1-5 years of the 
relationship) as can be seen in the last figure (bottom left pink), but also note 
that the relationship lengths that follow this pattern are mainly under 20 years 
in length and greater than 1 year. Another interesting pattern that can be noted 
is that for those who have been in a very long-term relationship (greater than 20 
years), the majority of those couples sleeping in separate beds every night or a 
few times per week started this sleeping situation within 16 to 20 year or more
than 20 years into their relationship. So, the couples with long relationships 
actually did not sleep separately until many years together had passed. 

The first two graphs do fit what the public might expect to see from a survey like 
this, however, it does not give any indication as to the reasons couples who have 
been together longer might be sleeping separately. The following sections and figures 
will aim to further examine this aspect of the population sample. More specifically,
could occupation and income (which are often tied to how busy or occupied a person 
might be in their daily activities) be related to couple sleeping separately? 
Additionally, what are the reasons for certain categories of couples (older, younger, etc) 
sleeping separately and have they noticed changes in their relationship from sleeping separately? 



***


## 2.2 Occupation and Income (Graph 2 and Graph 3) 

**Relate occupation, income, and bed status with how often they sleep separately.**

**Do occupational demands affect sleeping status? Is someone with a more demanding job more likely to sleep separately?**
**(Could call for inclusion or examination of another data-set that has information on occupational categories and average hours worked per day or week.)**


### 2.2.1 Graph 2
```{r 2b_graph2}
# GRAPH 2

# change the default theme below:
theme_set(theme_bw())
# set levels in order to give a better visisualization in color
#summary(sleeping_alone_data$How_Often_Sep_Beds)

rank<-as.factor(sleeping_alone_data$How_Often_Sep_Beds)
sleeping_alone_data4 <- sleeping_alone_data %>%
  mutate(How_Often_Sep_Beds = factor(rank, ordered = TRUE, 
                                     levels = c("Never",
                                               "Once a year or less",
                                               "Once a month or less",
                                               "A few times per month",
                                               "A few times per week",
                                               "Every night")))
# Relate occupation to bed status (and how often they sleep)
# Overall plot
sleeping_alone_data4 %>%
  group_by(Current_Occupation)%>%
  ggplot(data = sleeping_alone_data4,
       mapping = aes(x = Current_Occupation,
                     fill = How_Often_Sep_Beds)) +
  geom_bar(position="dodge")+
  
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.25,
                                   size = 7),
        legend.position = "top",
        legend.text = element_text(size = 7),
        legend.key.size = unit(0.25,'cm'),
        legend.title = element_text(size= 8),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

# summary() to see each category name and ready to draw plots for each of 
# the categories. facet_wrap or facet_grid provide too little information
#summary(sleeping_alone_data$How_Often_Sep_Beds)
#### for each bed status below:
# Never
sleeping_alone_data4 %>%
  select(Current_Occupation,How_Often_Sep_Beds) %>%
  filter(How_Often_Sep_Beds == "Never") %>%
  ggplot(aes(y = Current_Occupation,
             color = How_Often_Sep_Beds))+
  geom_bar(fill = "purple4")+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())
  
# Once a year or less 
sleeping_alone_data4 %>%
  select(Current_Occupation,How_Often_Sep_Beds) %>%
  filter(How_Often_Sep_Beds == "Once a year or less") %>%
  ggplot(aes(y = Current_Occupation,
             color = How_Often_Sep_Beds))+
  geom_bar(fill = "darkslateblue")+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())
# Once a month or less
sleeping_alone_data4 %>%
  select(Current_Occupation,How_Often_Sep_Beds) %>%
  filter(How_Often_Sep_Beds == "Once a month or less") %>%
  ggplot(aes(y = Current_Occupation,
             color = How_Often_Sep_Beds))+
  geom_bar(fill = "cadetblue4")+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())
# A few times per month 
sleeping_alone_data4 %>%
  select(Current_Occupation,How_Often_Sep_Beds) %>%
  filter(How_Often_Sep_Beds == "A few times per month") %>%
  ggplot(aes(y = Current_Occupation,
             color = How_Often_Sep_Beds))+
  geom_bar(fill = "lightseagreen")+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())
# A few times per week
sleeping_alone_data4 %>%
  select(Current_Occupation,How_Often_Sep_Beds) %>%
  filter(How_Often_Sep_Beds == "A few times per week") %>%
  ggplot(aes(y = Current_Occupation,
             color = How_Often_Sep_Beds))+
  geom_bar(fill = "palegreen3")+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())
# Every night
sleeping_alone_data4 %>%
  select(Current_Occupation,How_Often_Sep_Beds) %>%
  filter(How_Often_Sep_Beds == "Every night") %>%
  ggplot(aes(y = Current_Occupation,
              color= How_Often_Sep_Beds))+
  geom_bar(fill = "khaki1")+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())
## for people who have further response instead of checking boxes
## (choose to plot those who answered "retired" or "Retired)
sleeping_alone_data4 %>%
  filter(X.12 == "retired" | X.12 == "Retired")%>%
  mutate(X.12 ="retired") %>%
  
  ggplot(aes(x = X.12, fill= How_Often_Sep_Beds)) +
  
  geom_bar(color="white",position = "dodge")+
  
  theme(axis.text.x = element_text(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) 


```

### 2.2.2 Summarizing Paragraph:

People who are in Education, Training, and Library Occupations seems to have 
more response of "Never" sleeping separately. Regardless of the occupation,
The number of people who sleep separately is bigger than other frequencies.

For people who checked the box of "Other(please specify)", we select retired 
samples to draw a bar graph. Distribution of people who retired have heavy tails
on both side(Never & Every night), but the number of people who never sleeping 
separately is almost twice as large as those who sleeping separately every night.

***


### 2.2.3 Graph 3
```{r 2b_graph3}
# GRAPH 3

ggplot(data = sleeping_alone_data4,
       mapping = aes(x = Current_Occupation,
                     fill = How_Often_Sep_Beds)) +
  geom_bar()+
  scale_fill_viridis_d(option = "cividis")+
  
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.25,
                                   size = 7),
        legend.position = "top",
        legend.text = element_text(size = 7),
        legend.key.size = unit(0.25,'cm'),
        legend.title = element_text(size= 8),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())+
  facet_grid(~Gender)
## for each bed status below:
# Never
sleeping_alone_data4 %>%
  select(Current_Occupation,How_Often_Sep_Beds,Gender) %>%
  filter(How_Often_Sep_Beds == "Never") %>%
  ggplot(aes(x= Current_Occupation,
             color = How_Often_Sep_Beds,
             fill = Gender))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.25,
                                   size = 7),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())+
  facet_grid(~Gender)
  
# Once a year or less 
sleeping_alone_data4 %>%
  select(Current_Occupation,How_Often_Sep_Beds,Gender) %>%
  filter(How_Often_Sep_Beds == "Once a year or less") %>%
  ggplot(aes(x= Current_Occupation,
             color = How_Often_Sep_Beds,
             fill = Gender))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.25,
                                   size = 7),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())+
  facet_grid(~Gender)
# Once a month or less
sleeping_alone_data4 %>% filter(Gender != "") %>%
  select(Current_Occupation,How_Often_Sep_Beds,Gender) %>%
  filter(How_Often_Sep_Beds == "Once a month or less") %>%
  ggplot(aes(x= Current_Occupation,
             color = How_Often_Sep_Beds,
             fill = Gender))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.25,
                                   size = 7),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())+
  facet_grid(~Gender)
# A few times per month 
sleeping_alone_data4 %>% filter(Gender != "") %>%
  select(Current_Occupation,How_Often_Sep_Beds,Gender) %>%
  filter(How_Often_Sep_Beds == "A few times per month") %>%
  ggplot(aes(x= Current_Occupation,
             color = How_Often_Sep_Beds,
             fill = Gender))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.25,
                                   size = 7),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())+
  facet_grid(~Gender)
# A few times per week
sleeping_alone_data4 %>% filter(Gender != "") %>%
  select(Current_Occupation,How_Often_Sep_Beds,Gender) %>%
  filter(How_Often_Sep_Beds == "A few times per week") %>%
  ggplot(aes(x= Current_Occupation,
             color = How_Often_Sep_Beds,
             fill = Gender))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.25,
                                   size = 7),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())+
  facet_grid(~Gender)
# Every night
sleeping_alone_data4 %>% filter(Gender != "") %>%
  select(Current_Occupation,How_Often_Sep_Beds,Gender) %>%
  filter(How_Often_Sep_Beds == "Every night") %>%
  ggplot(aes(x= Current_Occupation,
             color = How_Often_Sep_Beds,
             fill = Gender))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.25,
                                   size = 7),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())+
  facet_grid(~Gender)


```

### 2.2.4 Summarizing Paragraph:

When we add gender as another consideration of our plots, it seems that there 
are more male participants for this study. For each frequencies of how often 
they sleep separately, there is a significant difference in number for the 
response of Never & Every night(males more than females). For other response 
of this question, the number in difference is not significant. People who
does not specify their gender has a large proportion of responding to "Never".

***



## 2.3 Reasons for Sleeping Separately (Graph 4) --> Matthew

**Relate how often couples sleep in different beds with the reason they sleep separately.**

**How correlated are the reasons they sleep in separate beds with the frequency they sleep together?**
**Reasons include snoring, frequent bathroom trips, sickness, intimacy issues, etc.**

### 2.3.1 Graph 4
```{r 2c_graph4}
# GRAPH 4

#Horizontal Bar Chart for Reason 1
sleeping_alone_data5 <- sleeping_alone_data %>% 
  
  group_by(Sep_Beds_Reasons, How_Often_Sep_Beds) %>% 
  
  summarise(freq = n()) %>% 
  
  mutate(Prop = round(freq/sum(freq), digits = 5)) %>% 
  
  select(Sep_Beds_Reasons, How_Often_Sep_Beds, Prop) %>% 
  
  slice(-c(1)) %>% 
  
  ungroup()
BarOftenVsReason1 <- ggplot(data = sleeping_alone_data5,
                               mapping = aes(y = Prop,
                                             x = How_Often_Sep_Beds,
                                             fill = How_Often_Sep_Beds)) +
  
  geom_bar(stat = "identity",
           show.legend = F) + 
  coord_flip() + 
  labs(title = "One of Us Snores",
       x = "",
       y = "") +
  scale_y_continuous(labels = scales::percent) +
  
  theme(plot.title = element_text(size = 15, hjust = 0.5))
#Horizontal Bar Chart for Reason 2
sleeping_alone_data6 <- sleeping_alone_data %>% 
  
  group_by(X.2, How_Often_Sep_Beds) %>% 
  
  summarise(freq = n()) %>% 
  
  mutate(Prop = round(freq/sum(freq), digits = 5)) %>% 
  
  select(X.2, How_Often_Sep_Beds, Prop) %>% 
  
  slice(-c(1)) %>% 
  
  ungroup()
BarOftenVsReason2 <- ggplot(data = sleeping_alone_data6,
                               mapping = aes(y = Prop,
                                             x = How_Often_Sep_Beds,
                                             fill = How_Often_Sep_Beds)) +
  
  geom_bar(stat = "identity",
           show.legend = F) + 
  coord_flip() + 
  labs(title = "Frequent Bathroom Trips",
       x = "",
       y = "") +
  
  scale_y_continuous(labels = scales::percent) +
  
  theme(plot.title = element_text(size = 15, hjust = 0.5))
#Horizontal Bar Chart for Reason 3
sleeping_alone_data7 <- sleeping_alone_data %>% 
  
  group_by(X.4, How_Often_Sep_Beds) %>% 
  
  summarise(freq = n()) %>% 
  
  mutate(Prop = round(freq/sum(freq), digits = 5)) %>% 
  
  select(X.4, How_Often_Sep_Beds, Prop) %>% 
  
  slice(-c(1)) %>% 
  
  ungroup()
BarOftenVsReason3 <- ggplot(data = sleeping_alone_data7,
                               mapping = aes(y = Prop,
                                             x = How_Often_Sep_Beds,
                                             fill = How_Often_Sep_Beds)) +
  
  geom_bar(stat = "identity",
           show.legend = F) + 
  coord_flip() + 
  labs(title = "No Longer Physically Intimate",
       x = "",
       y = "") +
  
  scale_y_continuous(labels = scales::percent) +
  
  theme(plot.title = element_text(size = 15, hjust = 0.5))
#Horizontal Bar Chart for Reason 4
sleeping_alone_data8 <- sleeping_alone_data %>% 
  
  group_by(X.5, How_Often_Sep_Beds) %>% 
  
  summarise(freq = n()) %>% 
  
  mutate(Prop = round(freq/sum(freq), digits = 5)) %>% 
  
  select(X.5, How_Often_Sep_Beds, Prop) %>% 
  
  slice(-c(1)) %>% 
  
  ungroup()
BarOftenVsReason4 <- ggplot(data = sleeping_alone_data8,
                               mapping = aes(y = Prop,
                                             x = How_Often_Sep_Beds,
                                             fill = How_Often_Sep_Beds)) +
  
  geom_bar(stat = "identity",
           show.legend = F) + 
  coord_flip() + 
  labs(title = "Different Temperature Preferences",
       x = "",
       y = "") +
  
  scale_y_continuous(labels = scales::percent) +
  
  theme(plot.title = element_text(size = 15, hjust = 0.5))
#Horizontal Bar Chart for Reason 5
sleeping_alone_data9 <- sleeping_alone_data %>% 
  
  group_by(X.6, How_Often_Sep_Beds) %>% 
  
  summarise(freq = n()) %>% 
  
  mutate(Prop = round(freq/sum(freq), digits = 5)) %>% 
  
  select(X.6, How_Often_Sep_Beds, Prop) %>% 
  
  slice(-c(1)) %>% 
  
  ungroup()
BarOftenVsReason5 <- ggplot(data = sleeping_alone_data9,
                               mapping = aes(y = Prop,
                                             x = How_Often_Sep_Beds,
                                             fill = How_Often_Sep_Beds)) +
  
  geom_bar(stat = "identity",
           show.legend = F) + 
  coord_flip() + 
  labs(title = "Arguments / Fights",
       x = "",
       y = "") +
  
  scale_y_continuous(labels = scales::percent) +
  
  theme(plot.title = element_text(size = 15, hjust = 0.5))
#Horizontal Bar Chart for Reason 6
sleeping_alone_data10 <- sleeping_alone_data %>% 
  
  group_by(X.8, How_Often_Sep_Beds) %>% 
  
  summarise(freq = n()) %>% 
  
  mutate(Prop = round(freq/sum(freq), digits = 5)) %>% 
  
  select(X.8, How_Often_Sep_Beds, Prop) %>% 
  
  slice(-c(1)) %>% 
  
  ungroup()
BarOftenVsReason6 <- ggplot(data = sleeping_alone_data10,
                               mapping = aes(y = Prop,
                                             x = How_Often_Sep_Beds,
                                             fill = How_Often_Sep_Beds)) +
  
  geom_bar(stat = "identity",
           show.legend = F) + 
  coord_flip() + 
  labs(title = "Not Wanting to Share Covers",
       x = "",
       y = "") +
  
  scale_y_continuous(labels = scales::percent) +
  
  theme(plot.title = element_text(size = 8, hjust = 0.5))
grid.arrange(BarOftenVsReason1, BarOftenVsReason2, BarOftenVsReason3, BarOftenVsReason4, BarOftenVsReason5, BarOftenVsReason6,nrow = 2,
             top=textGrob("Comparing How Often Couples Sleep In Different Beds With Reasons They Sleep Separately", 
             gp=gpar (fontsize=10, fontface = "bold")))

```


```{r 2c_graph4.1}
# GRAPH 5

help_stay_together <- sleeping_alone_data %>% filter(Statement_Help_Stay_Together != "") %>%
                      ggplot(mapping = aes(x = Statement_Help_Stay_Together,
                                           fill = How_Often_Sep_Beds)) +
                     geom_bar(position = "dodge") + 
                     labs(x = "Helped Us Stay Together",
                          fill = "How Often Couples Sleep Separately",
                          title = "Frequency Sleeping Separately and Improvements") +
                     theme(legend.position = "NA", 
                           title = element_text(size = 9)) +
                     coord_flip()

better_sleep <- sleeping_alone_data %>% filter(Statement_Better_Sleep != "") %>%
                      ggplot(mapping = aes(x = Statement_Better_Sleep,
                                           fill = How_Often_Sep_Beds)) +
                     geom_bar(position = "dodge") + 
                     labs(x = "Helped Improve Sleep",
                          fill = "How Often Couples Sleep Separately") +
                     theme(legend.position = "NA" ) +
                     coord_flip()

improved_sex_life <- sleeping_alone_data %>% filter(Statement_Improved_Sex_Life != "") %>%
                      ggplot(mapping = aes(x = Statement_Improved_Sex_Life,
                                           fill = How_Often_Sep_Beds)) +
                     geom_bar(position = "dodge") + 
                     labs(x = "Helped Improve Sex Life",
                          fill = "How Often Couples Sleep Separately") +
                     theme(legend.position = "bottom",
                           legend.key.size = unit(0.25, 'cm'), 
                           legend.title = element_blank(), 
                           legend.text = element_text(size=9)) +
                     coord_flip()



plot_grid(help_stay_together, better_sleep, improved_sex_life, ncol = 2, nrow = 2, label_size = 9) 

?plot_grid

```
### 2.3.2 Summarizing Paragraph:

These graphs compare how often couples sleep in different beds with some reasons 
why they sleep separately. The reason that the percentages add up to over 100% is 
because this part of the survey involved selecting all that applied. One specific 
pattern that I notice is that sleeping separately "a few times per month" is 
clearly the least common response among all reasons. Most of the reasons have the 
highest percentage of sleeping separately "every night", except for arguments and 
fights which has by far the lowest percentage. This makes sense to think about
because couples do not usually have arguments or fights every night. They generally 
happen on occasion, which is why the arguments/fights graph has the highest percentage 
of sleeping separately "once a year or less" and "once a month or less". Besides 
arguments and fights, the graphs have similar distributions.

In terms of imrpovements, majority of couples who sleep in separate beds every night 
who did not anser neutrally, reported strongly agreeing with the statement "it helped save our relationship". 
Most notably, the vast majority of couples sleeping separately every night stated that sleeping 
separately improved their sleep. When it comes to improvements in sex life, the results are 
less descriptive and there seems to be less unanimous improvements in this category. 
Overall, the reasons that couples chose to sleep apart every night were practical sleep reasons
relating to sleeping habits such as temperature differences, middle-of-the-night disturbances, 
snoring, and cover-sharing. 


***


# 3 Conclusions:  

This data does illustrate a tendancy toward stereotypical patterns in our society 
regarding age / relationship length and sleeping situation. That being that couples 
who are older (or have one partner who is older) are more likely to sleep in 
separate beds. Additionally couples who have been together for at least a decade are
much more likely to sleep in separate beds. That being said, the purpose of this data was
to help shed light on the factors involved in sleeping separately and allow us to examine 
the reasons why couples choose ot sleep separately. It seems that the reasons are highly
related to comfort and to getting the best quality of sleep. The stereotype of age / relationship
length and separate sleeping usually has to do with stereotypes about "romance fading away" or
things of the kind. However, the most popular reasons in general are far more specific and practical.
It is understandable that couples who have been together longer (say 16-20) years may 
be at the point where they can comfortably communicate that it would be best to have separate beds. 

# 4 Limitations / Recommendations: 

There are several limitations to this data. The aspect of gender identity is 
neglected in this survey, only male and female options are given and there is no
indication of which gender identity the partner or the person taking the survey is. 
It would be an interesting aspect to consider the patterns among different types of 
couples in the US (in terms of gender and sexual identity and how that varies depending on 
age/generation).






***







